```{r global_options, include=FALSE}
opts_chunk$set(fig.width=12, fig.height=8, echo=TRUE, warning=FALSE, message=FALSE);
```

Do it correctly
========================================================



<<<<<<< HEAD
## Introduction
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 
=======
#Exec summary

>>>>>>> 3c538e9c87c676219c6ba54e8f5b26ed29740e88

## Exploration of data

### Getting the data
The data are aquired from: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv (training set) and from https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv (testing set). Data are described here: http://groupware.les.inf.puc-rio.br/har

```{r}
library(ggplot2);
library(caret);

if (file.exists("pml-training.rds")) {
    pml_training<-readRDS("pml-training.rds");
} else {
    download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile="pml-training.csv", method="curl");
    pml_training<-read.csv("pml-training.csv", stringsAsFactors=FALSE, na.strings = c("NA", "#DIV/0!"));
    saveRDS(pml_training, file="pml-training.rds");
}

if (file.exists("pml-testing.rds")) {
    pml_testing<-readRDS("pml-testing.rds");
} else {
    download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile="pml-testing.csv", method="curl");
    pml_testing<-read.csv("pml-testing.csv", stringsAsFactors=FALSE, na.strings = c("NA", "#DIV/0!"));
    saveRDS(pml_testing, file="pml-testing.rds");
}

```
### Creation of a set
```{r}
set.seed(8888)
trainingIndex <- createDataPartition(pml_training$classe, list=FALSE, p=.7)
training = pml_training[trainingIndex,]
testing = pml_training[-trainingIndex,]
```

<<<<<<< HEAD
### Removal of non predicitve data

=======
>>>>>>> 3c538e9c87c676219c6ba54e8f5b26ed29740e88
Remove indicators with near zero variance. Those variables will not have a great impact on generated model.

```{r}
nzv <- nearZeroVar(training)

training <- training[-nzv]
testing <- testing[-nzv]
pml_testing <- pml_testing[-nzv]
```

Then we calculate which columns are numeric (since only those can be processed by algorithms)
```{r}
num_features_idx = which(lapply(training,class) %in% c('numeric')  )
```

We fill missing values based od model from `knnImpute` method from `preProcess` function.
```{r}
preModel <- preProcess(training[,num_features_idx], method=c('knnImpute'))

p_training <- cbind(training$classe, predict(preModel, training[,num_features_idx]))
p_testing <- cbind(testing$classe, predict(preModel, testing[,num_features_idx]))
pr_testing <- predict(preModel, pml_testing[,num_features_idx])

#Fixing names
names(p_training)[1] <- 'classe'
names(p_testing)[1] <- 'classe'
```

### Classification
Read more: http://groupware.les.inf.puc-rio.br/har#weight_lifting_exercises#ixzz38Q6Fxj5v
Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

```{r}
cvControl <- trainControl(method = "cv", number = 5)
treefit <- train(classe ~ ., trControl = cvControl, method = "rpart", tuneLength = 5, data=p_training);

plot(treefit)

# Due to time constraints, result of processing random forest with crossValidation are skipped, final result
# comes to model randomForest(classe ~ ., mtry=46, n=500)
if (file.exists("rForest.RData")) {
    load("rForest.RData")
} else {
    rForest <- train(classe ~ ., trControl = cvControl, method = "rf", tuneLength = 5, data=p_training);
    save(rForest, file="rForest.RData")
}
plot(rForest)

```

From the two plots, we can see that, using 5-fold cross validation, the estimated out of sample error rate of the Tree model is above 40%, while that of the Random Forest model is less than 1%. So we choose the best Random Forest model as our final model and apply it to our generated testing data to see how well it generalizes.

### Final model

```{r}
# model <- rForest$finalModel
library(randomForest)
model  <- randomForest(classe ~ ., p_training, ntree=500, mtry=46)
summary(model)
```

# Cross Validation
To measure the accuracy  training set and our cross validation set will be used. With the training set we can detect if our model has bias due to ridgity of model. With the cross validation set, determination if of variance due to overfitting can be detected.

##In-sample accuracy
```{r}
pred.train <- predict(model, newdata = p_training)
confusionMatrix(pred.train, p_training$classe)
```
The in sample accuracy is 100% which indicates, the model does not suffer from bias.

## Out-of-sample accuracy
```{r}
pred.test <- predict(model, newdata = p_testing)
confusionMatrix(pred.test, p_testing$classe)
```
The cross validation accuracy is greater than 99%, which should be sufficient for predicting the twenty test observations. Based on the lower bound of the confidence interval we would expect to achieve a 98.7% classification accuracy on new data provided.


## Answers to final testing model
```{r}
answers <- predict(model, pr_testing) 
print(answers)
```

<!--
https://github.com/dmaust/DataScience-ML-Project
https://github.com/SweeRoty/pml/blob/master/project.md

pml_write_files = function(x){
     n = length(x)
     for(i in 1:n){
         filename = paste0("problem_id_",i,".txt")
         write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
     }
 }
pml_write_files(answers)
-->
